%% -*- coding:utf-8 -*-
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   $RCSfile: grammatiktheorie-include.tex,v $
%%  $Revision: 1.13 $
%%      $Date: 2010/11/16 08:40:32 $
%%     Author: Stefan Mueller (CL Uni-Bremen)
%%    Purpose: 
%%   Language: LaTeX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



% http://tex.stackexchange.com/questions/38607/no-room-for-a-new-dimen
\usepackage{etex}\reserveinserts{28}

% http://tex.stackexchange.com/questions/229500/tikzmark-and-xelatex
% temporary fix, remove later
%\newcount\pdftexversion \pdftexversion140 \def\pgfsysdriver{pgfsys-dvipdfm.def} \usepackage{tikz} \usetikzlibrary{tikzmark}

%\usepackage[section]{placeins}

\usepackage{eurosym} % should go once Berthold fixes unicode

% http://tex.stackexchange.com/questions/284097/subscript-like-math-but-without-the-minus-sign?noredirect=1#comment685345_284097
% for subscripts
\usepackage{amsmath}
\usepackage{unicode-math} 
\newcommand\mathdash{\text{\normalfont -}}

% \justify to switch of \raggedright in translations
\usepackage{ragged2e}

% Haitao Liu
\usepackage{xeCJK}
\setCJKmainfont{SimSun}


\hypersetup{bookmarksopenlevel=0}

\iftoggle{draft}{
\usepackage{todonotes}
}{
\usepackage[disable]{todonotes}
}

\newcommand{\todostefan}[1]{\todo[color=green!40]{\footnotesize #1}\xspace}
\newcommand{\todosatz}[1]{\todo[color=red!40]{\footnotesize #1}\xspace}
\newcommand{\todoandrew}[1]{\todo[color=blue!40]{\footnotesize #1}\xspace}

\newcommand{\inlinetodostefan}[1]{\todo[color=green!40,inline]{\footnotesize #1}\xspace}
\newcommand{\inlinetodoandrew}[1]{\todo[color=red!40,inline]{\footnotesize #1}\xspace}


%% \newcommand{\remarkstefan}[1]{\todo[color=green!40]{\footnotesize #1}\xspace}
%% \newcommand{\remarkbjarne}[1]{\todo[color=red!40]{\footnotesize #1}\xspace}

%% \newcommand{\inlineremarkstefan}[1]{\todo[color=green!40,inline]{\footnotesize #1}\xspace}
%% \newcommand{\inlineremarkbjarne}[1]{\todo[color=red!40,inline]{\footnotesize #1}\xspace}

\newcommand{\treeag}{TAG\indextag}

%% is done by package option 
\ifdraft
\proofmodetrue
\fi


%% % taken from covington.sty (check)
%% %\newcounter{lsptempcnt}

%% \newcommand{\mex}[1]{\setcounter{lsptempcnt}{\value{equation}}%
%% \addtocounter{lsptempcnt}{#1}%
%% \arabic{lsptempcnt}}%


\usepackage{metalogo} % xelatex

\usepackage{multicol}

\usepackage{LSP/lsp-styles/lsp-forest-setup}

\usepackage{bookmark}

\forestset{
      terminus/.style={tier=word, for children={tier=tabular}, for tree={fit=band}, for descendants={no path, align=left, l sep=0pt}},
      sn edges original/.style={for tree={parent anchor=south, child anchor=north,align=center,base=top}},
      no path/.style={edge path={}},
      set me left/.style={calign with current edge, child anchor=north west, for parent={parent anchor=south west}},
}


\usepackage{my-ccg-ohne-colortbl}



%\displaywidowpenalty=10000\relax
%\predisplaypenalty=-200\relax



\usepackage{jambox}



\usepackage{german}\selectlanguage{USenglish}


\usepackage[final]{epsfig}
\usepackage{graphicx}


%\newcommand{\mod}{\textsc{mod}\xspace}  % wegen beamer.cls nicht in abbrev.sty

%\usepackage[figuresright]{rotating}


\usepackage{makros.2e,article-ex,additional-langsci-index-shortcuts,
eng-date,
lastpage,float,comment,my-theorems,soul,tabularx}

% loaded in macros.2e \usepackage[english]{varioref}
% do not stop and warn! This will be tested in the final version
\vrefwarning



\usepackage{ogonek}        % For Ewa Dabrowska

% http://tex.stackexchange.com/questions/203/how-to-obtain-verbatim-text-in-a-footnote
% somehow does not work
%\usepackage{fancyvrb}

%\newcommand{\tag}{TAG\indextag} % has to be here, conflict with latexbeamer

% mit der Index-Version geht die Silbentrennung nicht
\renewcommand{\word}[1]{\emph{#1}}

%\newcommand{\dom}{\textsc{dom}\xspace}

\newcommand{\prt}{\textsc{prt}}
%\newcommand{\refl}{\textsc{refl}}

%% \newcommand{\snom}{\textit{snom}}
%% \newcommand{\sgen}{\textit{sgen}}
%% \newcommand{\sacc}{\textit{sacc}}

%\usepackage{my-index-shortcuts}

\newcommand{\tes}{Tesnière\xspace}
\newcommand{\dom}{\textsc{dom}\xspace}


\newcommand{\page}{}

\usepackage{mycommands}% \dash

\let\mc=\multicolumn

% still needed
\usepackage{tikz-qtree}

\usepackage{LSP/lsp-styles/lsp-gb4e}
%\exewidth{\exnrfont (34)}
% should be set up for the whole series in langsci.cls
\renewcommand{\fnexfont}{\footnotesize\upshape}
%\let\oldglt\glt
%\def\glt{\oldglt\justify}
%\def\glt{\nopagebreak\vskip.17\baselineskip\transfont\parindent0ex}

\makeatletter
\def\ea{\ifnum\@xnumdepth=0\begin{exe}\else\begin{xlist}[iv.]\fi\ex}
\def\eal{\begin{exe}\exnrfont\ex\begin{xlist}[iv.]}

\def\gll%                  % Introduces 2-line text-and-gloss.
    {\bgroup %\begin{flushleft}
      \raggedright
     \ifx\@gsingle1%           conditionally force single spacing (hpk/MC)
	 \def\baselinestretch{1}\@selfnt\fi
%        \vskip\baselineskip\def\baselinestretch{1}%
%        \@selfnt\vskip-\baselineskip\fi%
    \bgroup
    \twosent
   }
\makeatother


\usepackage{subfig}

%\renewcommand{\xbar}{X̅\xspace}


% should be removed once that the \Tree figures are removed
\usepackage{forest}

\usepackage{pstricks,pst-node}

%\nodemargin5pt%\treelinewidth2pt\arrowwidth6pt\arrowlength10pt
\psset{nodesep=5pt} %,linewidth=0.8pt,arrowscale=2}
\psset{linewidth=0.5pt}
\setcounter{secnumdepth}{4}


\usepackage{dgmacros,pst-tree,trees,dalrymple} % Mary Dalrymples macros


% to set the MRSes for scope underspecification
%http://tex.stackexchange.com/questions/218417/replacing-tree-dvips-connect-nodes-in-a-tabular-environment/218458#218458
\usepackage{tcolorbox}
\tcbuselibrary{skins}
% for texlive 2015
\newtcbox{\mybox}[1][]{empty,shrink tight,nobeforeafter,on line,before upper=\vphantom{gM},remember as=#1,top=2pt,bottom=2pt}

% for texlive 2013
%\newtcbox{\mybox}[1][]{enhanced,boxrule=0pt,colframe=white,colback=white,shrink tight,nobeforeafter,on line,before upper=\vphantom{gM},remember as=#1} %,top=3pt,bottom=3pt}
                                %use shorten <=2pt,shorten >=2pt in the pictures.


\newcommand{\mynode}[2]{\mybox[#1]{#2}}


% http://tex.stackexchange.com/questions/218417/replacing-tree-dvips-connect-nodes-in-a-tabular-environment/218458#218458
% Instead of using the package tikzmark, you can define your own \tikzmark being a regular node. There's no need to use tcolorbox package.
\newcommand{\mysubnode}[2]%
    {\tikz[baseline=(#1.base), remember picture]\node[outer sep=0pt, inner sep=0pt] (#1) {#2};}


%%% trick for using adjustbox
\let\pstricksclipbox\clipbox
\let\clipbox\relax

% http://tex.stackexchange.com/questions/206728/aligning-several-forest-trees-in-centered-way/206731#206731
% for aligning TAG trees
\usepackage[export]{adjustbox}


% http://tex.stackexchange.com/questions/230300/doing-something-like-psframebox-in-tikz#230306
\tikzset{
frbox/.style={
  rounded corners,
  draw,
  thick,
  inner sep=5pt
  }
}
\newcommand\TZbox[1]{\tikz\node[frbox,baseline] {#1};}

% draw a grid for getting the coordinates
\usepackage{tikz-grid}

% for offsets in trees
\newlength{\offset}
\newlength{\offsetup}

\ifxetex
\usepackage{eng-hyp-utf8}
\else
\usepackage{eng-hyp}
\fi

\usepackage{appendix}

\input{localmetadata.tex}
\usepackage{localcommands}

\begin{document}
\renewcommand{\bibname}{References}

%\crop

\frontmatter
\maketitle

%\thispagestyle{empty}~\newpage\setcounter{page}{5}% for printing 2up

% Mit Widmung geht es auf S. 7 los
\thispagestyle{empty}~\newpage\setcounter{page}{7}% for printing 2up

%\pagestyle{scrheadings}%\ohead{\pagemark}\chead{}\ihead{\headmark}

\tableofcontents


\chapter*{Preface}

This book is an extended and revised version of my German book \emph{Grammatiktheorie}
\citep{MuellerGTBuch2}. It introduces various grammatical theories that play a role in current
theorizing or have made contributions in the past which are still relevant today. I explain some foundational
assumptions and then apply the respective theories to what can be called the ``core grammar'' of
German. I have decided to stick to the object language that I used in the German version of this
book since many of the phenomena that will be dealt with cannot be explained with English as the object
language. Furthermore, many theories have been developed by researchers with English as their native
language and it is illuminative to see these theories applied to another language.
I show how the theories under consideration deal with arguments and adjuncts, active/passive
alternations, local reorderings (so"=called scrambling), verb position, and fronting of phrases over
larger distances (the verb second property of the Germanic languages without English).

The second part deals with foundational questions that are important for developing theories.
This includes a discussion of the question of whether we have innate domain specific knowledge of
language (UG), the discussion of psycholinguistic evidence concerning the processing of language by
humans, a discussion of the status of empty elements and of the question whether we construct and perceive utterances 
holistically or rather compositionally, that is, whether we use phrasal or lexical constructions.

Unfortunately, linguistics is a scientific field 
with a considerable amount of terminological chaos. I therefore wrote an introductory
chapter that introduces terminology in the way it is used later on in the book. The second chapter
introduces phrase structure grammars, which plays a role for many of the theories that are covered
in this book. I use these two chapters (excluding the Section~\ref{sec-PSG-Semantik} on interleaving
phrase structure grammars and semantics) in introductory courses of our BA curriculum for German
studies. Advanced readers may skip these introductory chapters. The following chapters are
structured in a way that should make it possible to understand the introduction of the theories
without any prior knowledge. The sections regarding new developments and classification are more
ambitious: they refer to chapters still to come and also point to other publications that are
relevant in the current theoretical discussion but cannot be repeated or summarized in this
book. These parts of the book address advanced students and researchers. I use this book for teaching
the syntactic aspects of the theories in a seminar for advanced students in our BA. The slides are
available on my web page. The second part of the book, the general discussion, is more ambitious and contains the discussion
of advanced topics and current research literature.

This book only deals with relatively recent developments. For a historical overview, see for instance
\citew{Robins97a-u,JL2006a-u}. I am aware of the fact that chapters on
Integrational Linguistics\is{Integrational Linguistics}
\citep{Lieb83a-u,Eisenberg2004a,Nolda2007a-u}, Optimality Theory\indexot (\citealp{PS93a-u};
\citealp{Grimshaw97a-u}; G.\ \citealp{GMueller2000a-u}), Role and Reference Grammar\is{Role and
  Reference Grammar} \citep{vanValin93a-ed} and Relational Grammar\is{Relational Grammar}
\citep{Perlmutter83a-ed,Perlmutter84b-ed} are missing. I will leave these theories for later editions.

The original German book was planned to have 400 pages, but it finally was much bigger: the first
German edition has 525 pages and the second German edition has 564 pages. I
added a chapter on Dependency Grammar and one on Minimalism to the English version and now the
book has \pageref{LastPage} pages. I tried to represent the chosen theories appropriately and to cite all important work. Although the list of
references is over 85 pages long, I was probably not successful.
I apologize for this and any other shortcomings.

\section*{Acknowledgments}


I would like to thank David Adger\aimention{David Adger},
Jason Baldridge\aimention{Jason Baldridge}, 
Felix Bildhauer\aimention{Felix Bildhauer}, 
Emily M.\ Bender\aimention{Emily M. Bender},
Stefan Evert\aimention{Stefan Evert}, 
Gisbert Fanselow\aimention{Gisbert Fanselow}, 
Sandiway Fong\aimention{Sandiway Fong},
Hans"=Martin Gärtner\aimention{Hans"=Martin G"artner},
Kim Gerdes\aimention{Kim Gerdes},
Adele Goldberg\aimention{Adele Goldberg},
Bob Levine\aimention{Robert D. Levine},
Paul Kay\aimention{Paul Kay},
Jakob Maché\aimention{Jakob Mach{\'e}},
Guido Mensching\aimention{Guido Mensching},
Laura Michaelis\aimention{Laura Michaelis},
Geoffrey Pullum\aimention{Geoffrey K. Pullum}, 
Uli Sauerland\aimention{Uli Sauerland}, 
Roland Schäfer\aimention{Roland Sch"afer},
Jan Strunk\aimention{Jan Strunk},
Remi van Trijp\aimention{Remi van Trijp}, 
Shravan Vasishth\aimention{Shravan Vasishth},
Tom Wasow\aimention{Tom Wasow}, and
Stephen Wechsler\aimention{Stephen Mark Wechsler}
for discussion and 
%
Monika Budde\aimention{Monika Budde}, 
Philippa Cook\aimention{Philippa Helen Cook},
Laura Kallmeyer\aimention{Laura Kallmeyer}, 
Tibor Kiss\aimention{Tibor Kiss},
Gisela Klann"=Delius\aimention{Gisela Klann"=Delius}, 
Jonas Kuhn\aimention{Jonas Kuhn},
Timm Lichte\aimention{Timm Lichte}, % für Kommentare zum TAG"=Kapitel
Anke Lüdeling\aimention{Anke L"udeling},
Jens Michaelis\aimention{Jens Michaelis},
Bjarne Ørsnes\aimention{Bjarne {\O}rsnes},
Andreas Pankau\aimention{Andreas Pankau},     % Chomsky 2013
Christian Pietsch\aimention{Christian Pietsch},
Frank Richter\aimention{Frank Richter},
Ivan Sag\aimention{Ivan A. Sag}, and
Eva Wittenberg\aimention{Eva Wittenberg}
for comments on earlier versions of the German edition of this book and
%
%
Thomas Groß\aimention{Thomas M. Gro{\ss}},
Dick Hudson\aimention{Richard Hudson},
Sylvain Kahane\aimention{Sylvain Kahane}, 
Paul Kay\aimention{Paul Kay},
Haitao Liu (刘 海涛)\aimention{Haitao Liu},
Andrew McIntyre\aimention{Andrew McIntyre},
Sebastian Nordhoff\aimention{Sebastian Nordhoff},
Tim Osborne\aimention{Timothy Osborne}, 
Andreas Pankau\aimention{Andreas Pankau}, and
Christoph Schwarze\aimention{Christoph Schwarze}
for comments on earlier versions of this book. Thanks to Leonardo Boiko and Sven Verdoolaege for pointing out typos.
Special thanks go to Martin Haspelmath\aimention{Martin Haspelmath} for very detailed comments on an
earlier version of the English book. 

This book was the first Language Science Press book that had an open review phase (see below). I
thank Dick Hudson, Paul Kay, Antonio Machicao Y Priemer, Andrew McIntyre, Sebastian Nordhoff, and one anonymous open
reviewer for their comments. Theses comments are documented at the \href{\lsURL}{download page of this book}.



I thank Wolfgang Sternefeld and Frank Richter, who wrote a detailed review of the German version of
this book \citep{SR2012a}. They pointed out some mistakes and omissions that were corrected in the second edition
of the German book and which are of course not present in the English version.

Thanks to all the students who commented on the book and whose questions lead to improvements. 
Lisa Deringer,
Aleksandra Gabryszak, % Student SS 2010 gute Fragen, GB-Verbbewegung und LMT
Simon Lohmiller, %Student, Typos und allgemeine Anregung zu Einführungskapitel
Theresa Kallenbach, %Studentin, GPSG
Steffen Neu\-schulz,  % Student SS 2010 gute Fragen
Reka Meszaros-Segner,
Lena Terhart and
Elodie Winckel deserve special mention.

Since this book is built upon all my experience in the area of grammatical theory, I want to thank
all those with whom I ever discussed linguistics during and after talks at conferences, workshops,
summer schools or via email.
Werner Abraham,
John Bateman,
Dorothee Beermann, 
Rens Bod, 
Miriam Butt,
Manfred Bierwisch,
Ann Copestake, 
Holger Diessel, 
Kerstin Fischer,
Dan Flickinger,
Peter Gallmann, 
%Adele Goldberg,  included above
Petter Haugereid,
Lars Hellan, 
% Paul Kay, included above
Tibor Kiss,
Wolfgang Klein, 
Hans"=Ulrich Krieger,
%Emily M. Bender, included above
Andrew McIntyre,
Detmar Meurers,
%Laura Michaelis,  included above
Gereon Müller,
Martin Neef,
Manfred Sailer, 
Anatol Stefanowitsch,
Peter Svenonius,
Michael Tomasello, 
Hans Uszkoreit, 
Gert Webelhuth, 
% Stephen Wechsler included above
Daniel Wiechmann and Arne Zeschel deserve special mention.

I thank Sebastian Nordhoff for a comment regarding the completion of the subject index entry for \emph{recursion}\is{recursion}.

Andrew Murphy translated part of Chapter~1 and the Chapters~2--3, 5--10, and 12--23. Many thanks for this!

During the past years, a number of workshops on theory comparison have taken place. I was invited to three of them.
I thank Helge Dyvik\aimention{Helge Dyvik} and Torbjørn Nordgård\aimention{Torbj{\o}rn
  Nordg{\r{a}}rd}\todostefan{Indexeinträge für Torbjorn and Bjarne do not work} for inviting me to the fall school for Norwegian PhD
students  \emph{Languages and Theories in Contrast}, which took place 2005 in Bergen. Guido Mensching\aimention{Guido Mensching} and Elisabeth
Stark\aimention{Elisabeth Stark} invited me to the workshop \emph{Comparing Languages and Comparing Theories:
  Generative Grammar and Construction Grammar}, which took place in 2007 at the Freie Universität
Berlin and Andreas Pankau\aimention{Andreas Pankau} invited me to the workshop \emph{Comparing
  Frameworks} in 2009 in Utrecht. I really enjoyed the discussion with all participants of these
events and this book benefited enormously from the interchange.

I thank Peter Gallmann\aimention{Peter Gallmann} for the discussion of his lecture notes on \gb
during my time in Jena. The Sections~\ref{Abschnitt-T-Modell}--\ref{Abschnitt-GB-Passiv} have a
structure that is similar to the one of his script and take over a lot. Thanks to David Reitter for
the \LaTeX{} macros for Combinatorial Categorial Grammar, to Mary Dalrymple and Jonas Kuhn for the LFG
macros and example structures, and to Laura Kallmeyer for the \LaTeX{} sources of most of the TAG
analyses. Most of the trees have been adapted to the \texttt{forest} package because of compatibility issues
with \XeLaTeX, but the original trees and texts were a great source of inspiration and without them
the figures in the respective chapters would not be half as pretty as they are now.

I thank Sašo Živanović for implementing the \LaTeX{} package \texttt{forest}. It really simplifies
typesetting of trees, dependency graphs, and type hierarchies. I also thank him for individual help
via email and on \href{http://www.stackexchange.com}{stackexchange}. In general, those active on stackexchange could not be thanked
enough: most of my questions regarding specific details of the typesetting of this book or the
implementation of the \LaTeX{} classes that are used by \lsp now have been answered within several
minutes. Thank you! Since this book is a true open access book under the CC-BY license, it can also
be an open source book. The interested reader finds a copy of the source code at \textbf{specify
  url}.\todostefan{specify URL} By making the book open source I pass on the knowledge provided by the \LaTeX{} gurus and
hope that others benefit from this and learn to typeset their linguistics papers in nicer and/or
more efficient ways.


Viola Auermann and Antje Bahlke, Sarah Dietzfelbinger, Lea Helmers, and Chiara Jancke cannot be thanked enough for their work at the copy machines. Viola
also helped a lot with proof reading prefinal stages of the translation.
I also want to thank my (former) lab members Felix Bildhauer, Philippa Cook, Janna Lipenkova, Jakob Maché,
Bjarne Ørsnes and Roland Schäfer\aimention{Roland Sch{\"a}fer}, which were mentioned above already
for other reasons, for their help with teaching. During the years from 2007 until the publication of
the first German edition of this book two of the three tenured positions in German Linguistics were
unfilled and I would have not been able to maintain the teaching requirements without their help and
would have never finished the \emph{Grammatiktheorie} book.

I thank Tibor Kiss for advice in questions of style. His diplomatic way always was a shining
example for me and I hope that this is also reflected in this book.


\section*{On the way this book is published}

I started to work on my dissertation in 1994 and defended it in 1997. During the whole time the
manuscript was available on my web page. After the defense, I had to look for a publisher. I was
quite happy to be accepted to the series \emph{Linguistische Arbeiten} by Niemeyer, but at the same time I
was shocked about the price, which was 186.00 DM for a paperback book that was written and typeset
by me without any help by the publisher (twenty times the price of a paperback novel).\footnote{
  As a side remark: in the meantime Niemeyer was bought by de Gruyter and closed down. The price of the book is now
  139.95 \euro / \$ 196.00. The price in Euro corresponds to 273.72 DM. 
%This is a price increase of 47\,\%.
} This
basically meant that my book was depublished: until 1998 it was available from my web page and after
this it was available in libraries only. My Habilitationsschrift was published by CSLI Publications
for a much more reasonable price. When I started writing textbooks, I was looking for alternative
distribution channels and started to negotiate with no-name print on demand publishers. Brigitte Narr,
who runs the Stauffenburg publishing house, convinced me to publish my HPSG textbook with her. The
copyrights for the German version of the book remained with me so that I could publish it on my web page. The collaboration was successful so that I also published my second textbook about
grammatical theory with Stauffenburg. I think that this book has a broader relevance and should be
accessible for non-German-speaking readers as well. I therefore decided to have it translated into
English. Since Stauffenburg is focused on books in German, I had to look for another publisher. Fortunately the situation in the publishing sector changed quite dramatically in comparison
to 1997: we now have high profile publishers with strict peer review that are entirely open access. I am very
glad about the fact that Brigitte Narr sold the rights of my book back to me and that I can now 
publish the English version with Language Science Press under a CC-BY license.


\section*{Language Science Press: scholar-owned high quality linguistic books}

In 2012 a group of people found the situation in the publishing business so unbearable that they
agreed that it would be worthwhile to start a bigger initiative for publishing linguistics books in
platinum open access, that is, free for both readers and authors. I set up a web page and collected
supporters, very prominent linguists from all over the world and all subdisciplines and Martin
Haspelmath and I then founded Language Science Press. At about the same time the DFG had announced
a program for open access monographs and we applied \citep{MH2013a} and got funded (two out of 18 applications got
funding). The money is used for a coordinator (Dr.\ Sebastian Nordhoff) and an economist (Debora
Siller), two programmers (Carola Fanselow and Dr.\ Mathias Schenner), who work on the publishing
plattform Open Monograph Press (OMP) and on conversion software that produces various formats (ePub, XML,
HTML) from our \LaTeX{} code. Svantje Lilienthal works on the documentation of OMP, produces
screencasts and does user support for authors, readers and series editors.

OMP is extended by open review facilities and community-building gamification tools
\citep{MuellerOA,MH2013a}. All Language Science Press books are reviewed by at least two external
reviewers. Reviewers and authors may agree to publish these reviews and thereby make the whole
process more transparent (see also \citew{Pullum84a} for the suggestion of open reviewing of journal
articles). In addition there is an optional second review phase: the open
review. This review is completely open to everybody. The whole community may comment on the document
that is published by Language Science Press. After this second review phase, which usually lasts for
two months, authors may revise their publication and an improved version will be published. This
book was the first book to go through this open review phase. The annotated open review version of this book is still available via
the \href{\lsURL}{web page of this book}. 

Currently, Language Science Press has 16 series on various subfields of linguistics with high
profile series editors from all continents. Series editors  and authors are responsible for
delivering manuscripts that are typeset in \LaTeX{}, but they are supported by a web-based typesetting
infrastructure that was set up by Language Science Press and by volunteer typesetters from the
community. Proofreading is also community-based.

If you think that textbooks like this one should be freely available to whoever wants to read them
and that publishing scientific results should not be left to profit-oriented publishers, then you
can join the Language Science Press community and support us in various ways: you can register with Language Science Press and have your name
listed on our supporter page with more than 500 other enthusiasts, you may devote your time and help
with proofreading and/or typesetting, or you may donate money for specific books or for Language
Science Press in general. We are also looking for institutional supporters like foundations,
societies, linguistics departments or university libraries. In case of questions, please contact me
or the Language Science Press coordinator at \href{mailto:contact@langsci-press.org}{contact@langsci-press.org}.


~\medskip

\noindent
Berlin, \today\hfill Stefan Müller



\mainmatter

\part{Background and specific theories}

\include{1-begriffe}
\include{2-psg}
\include{3-gb}
\include{3-minimalism}
\include{4-gpsg}
\include{5-merkmalstrukturen}
\include{6-lfg}
\include{7-cg}
\include{8-hpsg}
\include{9-cxg}
\include{dg}
\include{tag}
%\include{ot}

\part{General discussion}
\label{part-discussion}

\include{innateness}
\include{mts}
\include{competence}
\include{acquisition}
\include{complexity}
%\include{branching}
%\include{locality}
%\include{recursion}
\include{branching-locality-recursion}
\include{empty}
\include{movement}
\include{phrasal}
\include{coregram}
\include{conclusions}

%\renewcommand{\appendixpagename}{Appendix}
%\renewcommand{\appendixtocname}{Appendix}
\appendix
%\appendixpage
%\part{Appendix}
\include{loesungen}

\input{gt-index-cross}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%             Backmatter                       %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
\bookmarksetup{startatroot}

\bibliography{bib-abbr,biblio,lfg-implementations,crossrefs,temp}
\clearpage

\small
   

\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{chapter}{Index} 
\addcontentsline{toc}{section}{Name index}
\ohead{Name index}
%with biblatex
%\printindex
%without it 
\printindex 
  
\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Language index}
\ohead{Language index} 
\printindex[lan] 
  
\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Subject index}
\ohead{Subject index}
\printindex[sbj]


\end{document} 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%                  END                         %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%





%      <!-- Local IspellDict: en_US-w_accents -->
