\section*{Acknowledgments}


I would like to thank David Adger\aimention{David Adger},
Jason Baldridge\aimention{Jason Baldridge}, 
Felix Bildhauer\aimention{Felix Bildhauer}, 
Emily M.\ Bender\aimention{Emily M. Bender},
Stefan Evert\aimention{Stefan Evert}, 
Gisbert Fanselow\aimention{Gisbert Fanselow}, 
Sandiway Fong\aimention{Sandiway Fong},
Hans"=Martin Gärtner\aimention{Hans"=Martin G"artner},
Kim Gerdes\aimention{Kim Gerdes},
Adele Goldberg\aimention{Adele Goldberg},
Bob Levine\aimention{Robert D. Levine},
Paul Kay\aimention{Paul Kay},
Jakob Maché\aimention{Jakob Mach{\'e}},
Guido Mensching\aimention{Guido Mensching},
Laura Michaelis\aimention{Laura Michaelis},
Geoffrey Pullum\aimention{Geoffrey K. Pullum}, 
Uli Sauerland\aimention{Uli Sauerland}, 
Roland Schäfer\aimention{Roland Sch"afer},
Jan Strunk\aimention{Jan Strunk},
Remi van Trijp\aimention{Remi van Trijp}, 
Shravan Vasishth\aimention{Shravan Vasishth},
Tom Wasow\aimention{Tom Wasow}, and
Stephen Wechsler\aimention{Stephen Mark Wechsler}
for discussion and 
%
Monika Budde\aimention{Monika Budde}, 
Philippa Cook\aimention{Philippa Helen Cook},
Laura Kallmeyer\aimention{Laura Kallmeyer}, 
Tibor Kiss\aimention{Tibor Kiss},
Gisela Klann"=Delius\aimention{Gisela Klann"=Delius}, 
Jonas Kuhn\aimention{Jonas Kuhn},
Timm Lichte\aimention{Timm Lichte}, % für Kommentare zum TAG"=Kapitel
Anke Lüdeling\aimention{Anke L"udeling},
Jens Michaelis\aimention{Jens Michaelis},
Bjarne Ørsnes\aimention{Bjarne {\O}rsnes},
Andreas Pankau\aimention{Andreas Pankau},     % Chomsky 2013
Christian Pietsch\aimention{Christian Pietsch},
Frank Richter\aimention{Frank Richter},
Ivan Sag\aimention{Ivan A. Sag}, and
Eva Wittenberg\aimention{Eva Wittenberg}
for comments on earlier versions of the German edition of this book and
%
%
Thomas Groß\aimention{Thomas M. Gro{\ss}},
Dick Hudson\aimention{Richard Hudson},
Sylvain Kahane\aimention{Sylvain Kahane}, 
Paul Kay\aimention{Paul Kay},
Haitao Liu (刘 海涛)\aimention{Haitao Liu},
Andrew McIntyre\aimention{Andrew McIntyre},
Sebastian Nordhoff\aimention{Sebastian Nordhoff},
Tim Osborne\aimention{Timothy Osborne}, 
Andreas Pankau\aimention{Andreas Pankau}, and
Christoph Schwarze\aimention{Christoph Schwarze}
for comments on earlier versions of this book. Thanks to Leonardo Boiko and Sven Verdoolaege for pointing out typos.
Special thanks go to Martin Haspelmath\aimention{Martin Haspelmath} for very detailed comments on an
earlier version of the English book. 

This book was the first Language Science Press book that had an open review phase (see below). I
thank Dick Hudson, Paul Kay, Antonio Machicao y Priemer, Andrew McIntyre, Sebastian Nordhoff, and one anonymous open
reviewer for their comments. Theses comments are documented at the \href{\lsURL}{download page of
  this book}. In addition the book went through a stage of community proofreading (see also
below). Some of the proofreaders did much more than proofreading, their comments are highly
appreciated and I decided to publish these comments as additional open reviews.
Armin Buch, 
Leonel de Alencar,
Andreas Hölzl,
Gianina Iordăchioaia,
Timm Lichte,
Antonio Machicao y Priemer, and
Neal Whitman
deserve special mention here.


I thank Wolfgang Sternefeld and Frank Richter, who wrote a detailed review of the German version of
this book \citep{SR2012a}. They pointed out some mistakes and omissions that were corrected in the second edition
of the German book and which are of course not present in the English version.

Thanks to all the students who commented on the book and whose questions lead to improvements. 
Lisa Deringer,
Aleksandra Gabryszak, % Student SS 2010 gute Fragen, GB-Verbbewegung und LMT
Simon Lohmiller, %Student, Typos und allgemeine Anregung zu Einführungskapitel
Theresa Kallenbach, %Studentin, GPSG
Steffen Neu\-schulz,  % Student SS 2010 gute Fragen
Reka Meszaros-Segner,
Lena Terhart and
Elodie Winckel deserve special mention.

Since this book is built upon all my experience in the area of grammatical theory, I want to thank
all those with whom I ever discussed linguistics during and after talks at conferences, workshops,
summer schools or via email.
Werner Abraham,
John Bateman,
Dorothee Beermann, 
Rens Bod, 
Miriam Butt,
Manfred Bierwisch,
Ann Copestake, 
Holger Diessel, 
Kerstin Fischer,
Dan Flickinger,
Peter Gallmann, 
%Adele Goldberg,  included above
Petter Haugereid,
Lars Hellan, 
% Paul Kay, included above
Tibor Kiss,
Wolfgang Klein, 
Hans"=Ulrich Krieger,
%Emily M. Bender, included above
Andrew McIntyre,
Detmar Meurers,
%Laura Michaelis,  included above
Gereon Müller,
Martin Neef,
Manfred Sailer, 
Anatol Stefanowitsch,
Peter Svenonius,
Michael Tomasello, 
Hans Uszkoreit, 
Gert Webelhuth, 
% Stephen Wechsler included above
Daniel Wiechmann and Arne Zeschel deserve special mention.

I thank Sebastian Nordhoff for a comment regarding the completion of the subject index entry for \emph{recursion}\is{recursion}.

Andrew Murphy translated part of Chapter~1 and the Chapters~2--3, 5--10, and 12--23. Many thanks for this!

I also want to thank the 27 community proofreaders (\makeatletter\@proofreader\makeatother) that each worked on one or more chapters and
really improved this book. I got more comments from every one of them than I ever got for a book
done with a commercial publisher. Some comments were on content rather than on typos and layout
issues. No proofreader employed by a commercial publisher would have spotted these mistakes and
inconsistencies since commercial publishers do not have staff that knows all the grammatical
theories that are covered in this book. 

During the past years, a number of workshops on theory comparison have taken place. I was invited to three of them.
I thank Helge Dyvik\aimention{Helge Dyvik} and Torbjørn Nordgård\aimention{Torbj{\o}rn
  Nordg{\r{a}}rd}\todostefan{Indexeinträge für Torbjorn and Bjarne do not work} for inviting me to the fall school for Norwegian PhD
students  \emph{Languages and Theories in Contrast}, which took place 2005 in Bergen. Guido Mensching\aimention{Guido Mensching} and Elisabeth
Stark\aimention{Elisabeth Stark} invited me to the workshop \emph{Comparing Languages and Comparing Theories:
  Generative Grammar and Construction Grammar}, which took place in 2007 at the Freie Universität
Berlin and Andreas Pankau\aimention{Andreas Pankau} invited me to the workshop \emph{Comparing
  Frameworks} in 2009 in Utrecht. I really enjoyed the discussion with all participants of these
events and this book benefited enormously from the interchange.

I thank Peter Gallmann\aimention{Peter Gallmann} for the discussion of his lecture notes on \gb
during my time in Jena. The Sections~\ref{Abschnitt-T-Modell}--\ref{Abschnitt-GB-Passiv} have a
structure that is similar to the one of his script and take over a lot. Thanks to David Reitter for
the \LaTeX{} macros for Combinatorial Categorial Grammar, to Mary Dalrymple and Jonas Kuhn for the LFG
macros and example structures, and to Laura Kallmeyer for the \LaTeX{} sources of most of the TAG
analyses. Most of the trees have been adapted to the \texttt{forest} package because of compatibility issues
with \XeLaTeX, but the original trees and texts were a great source of inspiration and without them
the figures in the respective chapters would not be half as pretty as they are now.

I thank Sašo Živanović for implementing the \LaTeX{} package \texttt{forest}. It really simplifies
typesetting of trees, dependency graphs, and type hierarchies. I also thank him for individual help
via email and on \href{http://www.stackexchange.com}{stackexchange}. In general, those active on stackexchange could not be thanked
enough: most of my questions regarding specific details of the typesetting of this book or the
implementation of the \LaTeX{} classes that are used by \lsp now have been answered within several
minutes. Thank you! Since this book is a true open access book under the CC-BY license, it can also
be an open source book. The interested reader finds a copy of the source code at \url{https://github.com/langsci/25}. By making the book open source I pass on the knowledge provided by the \LaTeX{} gurus and
hope that others benefit from this and learn to typeset their linguistics papers in nicer and/or
more efficient ways.


Viola Auermann and Antje Bahlke, Sarah Dietzfelbinger, Lea Helmers, and Chiara Jancke cannot be thanked enough for their work at the copy machines. Viola
also helped a lot with proof reading prefinal stages of the translation.
I also want to thank my (former) lab members Felix Bildhauer, Philippa Cook, Janna Lipenkova, Jakob Maché,
Bjarne Ørsnes and Roland Schäfer\aimention{Roland Sch{\"a}fer}, which were mentioned above already
for other reasons, for their help with teaching. During the years from 2007 until the publication of
the first German edition of this book two of the three tenured positions in German Linguistics were
unfilled and I would have not been able to maintain the teaching requirements without their help and
would have never finished the \emph{Grammatiktheorie} book.

I thank Tibor Kiss for advice in questions of style. His diplomatic way always was a shining
example for me and I hope that this is also reflected in this book.

%      <!-- Local IspellDict: en_US-w_accents -->
