%% -*- coding:utf-8 -*-

\section*{Foreword of the fourth edition}



% reference to Sag2020

% We thank Shalom Lappin and Richard Sproat for discussion of implementation issues.

% added footnote to gb chapter regarding the assignment of semantic role accross phrase boundary

% Thanks Andreas Pankau

% trincated English -> truncated English 19.01.2020

% fixed URLs, added DOIs

% e or t for trace -> t for trace

% be- und ent-laden sind keine Partikel sondern Präfixe
% Namen ersetzt und Frauen durch Eichhörnchen

% added references for DP approach (Brame, Hudson)

% Due to the work on Chinese, some index entries were fixed (Phenomenon/phenomenon)

I fixed several typos, added and updated URLs and DOIs in the book and in the list of references.
I added a footnote to Chapter~\ref{chap-GB} concerning the assignment of semantic roles
across phrase boundaries (footnote~\ref{fn-semantic-role-phrase-boundary} on
p.\,\pageref{fn-semantic-role-phrase-boundary}). I thank Andreas Pankau for discussion on this point.

I added a paragraph discussing John Torr's implementational work (pages~\pageref{page-torr-implementation-beginning}--\pageref{page-torr-implementation-end}). I thank Shalom Lappin and Richard Sproat for discussion of implementation issues.

A small paragraph for further reading was added to Chapter~\ref{chap-phrasal} on phrasal vs.\ lexical analyses.

Language Science Press will publish a handbook on Head-Driven Phrase Structure Grammar hopefully
later this year \citep*{HPSGHandbook}. It contains several chapters comparing other syntactic
theories to HPSG. I added the respective references to the further readings sections of the chapters
for Lexical Functional Grammar, Categorial Grammar, Construction Grammar, and Minimalism.

This edition is the first edition that uses precompiled trees. Setting this up was not
straightforward. I am really grateful to Sašo Živanović for helping me and adapting the
\texttt{forest} package so that everything runs smoothly and efficiently. This saves me a lot of
time and reduces the energy consumption of my computer dramatically.


~\medskip

\noindent
Berlin, 2nd September 2020\hfill Stefan Müller



%      <!-- Local IspellDict: en_US-w_accents -->
